### 2020-03-24 ChapterProblemDefinition/chapterProblemDefinition.tex
Updated: Problem Definition summary renamed
Updated: Total proof-read and conclusion changed to approximate solution

\subsection{Object duplication or loss}
Any solution in which objects are migrated between servers is vulnerable to object duplication or loss. Networks do not provide absolute guarantee of message deliveries or order of message arrival. As a consequence of this, network messages for the migration of objects may be lost, in which case the object could be lost. For example, if one server sends a migrate message for an object and deletes the object from its simulation and the message is lost and the object never gets received and created on the receiving server. Alternatively, a server may not delete the migrating object until an acknowledgement is received from the server it is migrating the object to, in which case, the acknowledgement may be lost and the object kept on the sending server, despite a duplicate of the object now existing on the receiving server. %Messages may arrive out of order

@@

Accomplishing a transfer with a simple message from one server to another is charged with many problems: (1) the message may be lost and the object disappears; (2) the message may be delayed and the object disappears and reappears; (3) there may be more than one candidate server for hosting (if object travelling quickly) resulting in duplication of an object; (4) the server an object leaves may be unaware of when it should stop simulating the object (resulting in duplicated objects)\cite{High-performanceServerSystems}; (5) an object may be colliding with another object across the boundary and the collision would be missed if both objects are hosted on different servers (objects may pass through each other).

@@

There would be a trade-off with this approach, either:
1) Frequent messages are sent with small corrections for objects, appearing as jitter. This would require lots of computational and network overhead. 
2) Infrequent messages with large corrections for objects, appearing as objects moving large distances instantaneously but would require less computational and network overhead.

@@

This approach would be affected by network delays. For example, using the stacked cubes example. Say the two cubes are free-falling. If the bottom cube was being simulated on a different server, the position of the bottom cube would only be updated whenever a message was received from the remote server to do so. If there was a long delay between messages from the remote server, for example, from a spike in latency or a sudden big loss in packets, the cube would not be updated and appear to have stopped falling. As the bottom cube has stopped on the receiving server, the top cube remains stationary, as it is in contact with the bottom cube. On the remote server, if messages are being received, the cube would be stationary and appear to be floating, as from its server's perspective, it is in contact with the bottom cube. As a result, both servers have incorrect states compared to a centralised simulation.

\subsection{All object interactions on same server}
Simulate all objects that are interacting with each other on the same server. Auras could be used to determine which objects are potentially interacting. Auras of objects would need to be exchanged. Solution would need to manage the number of auras exchanged to reduce this overhead.

@@

In AP, an object maintains an aura, an area of interest around the object, that indicates its possible future location and, therefore, aids in identifying near future interactions. In other words, auras act as a request for objects to be migrated. Any object colliding with the aura is migrated to the server the aura came from. The presence of an aura allows: (1) the prediction of future hosting requirements allowing time to transfer objects; (2) a narrowing of interest in only considering a subset of objects promoting a scalable solution; (3) an unhindered physics simulation for existing objects; (4) limiting communications requirements between servers based on focusing message passing overhead on interacting objects (promoting scalability). In essence, AP ensures that any two objects that may be interacting are always being simulated on the same server. For example, objects C and D illustrated in Fig. \ref{AuraProj}. Object C projects its aura into Server 1, Object D collides with the aura and is migrated to Server 0, allowing continued interaction to occur.

@@

\section{Aura Calculation}
Calculating an aura size is key to creating a scalable solution balanced against consistency: (1) Auras that are too small will result in a faster simulation but with more missed interactions; (2) Auras that are too large will include more calculations that are simply not required, increasing server load and decreasing scalability. Therefore, how auras are calculated will be carefully described and justified.

@@

The solution to this problem is to dynamically move the server-region boundaries to prevent clusters of objects having the majority of the objects outside of their host server region or even avoid be moved to avoid intersecting large clusters of objects entirely. However, this solution is not yet implemented in AP.

@@

\section{Summary}
In this chapter, the following points have been addressed:
\begin{itemize}
	\item The requirements of a solution to scalable real-time physics.
	\item The challenges of distributing a real-time physics engine and potential solutions.
	\item A description and justification of AP, including the aura calculation.
	\item The implemented solutions to the problems of: 3-object thrashing; islands; and the corner case.
	\item The proposed solutions to: sub-optimal object hosting and 2-object thrashing, which are not yet implemented.
\end{itemize}

>>>>

\subsection{Object duplication or loss}
Any solution in which objects are migrated between servers is vulnerable to object duplication or loss. Networks do not provide absolute guarantee of message deliveries or order of message arrival. As a consequence of this, network messages for the migration of objects may be lost, in which case the object could be lost. For example, if one server sends a migrate message for an object and deletes the object from its simulation and the message is lost and the object never gets received and created on the receiving server. Alternatively, a server may not delete the migrating object until an acknowledgement is received from the server it is migrating the object to. In this case, the acknowledgement may be lost and the object kept on the sending server, despite a duplicate of the object now existing on the receiving server. %Messages may arrive out of order

@@

Accomplishing a transfer with a simple message from one server to another is charged with many problems: 
\begin{itemize}
	\item The message may be lost and the object disappears.
	\item The message may be delayed and the object disappears and reappears.
	\item There may be more than one candidate server for hosting (if object travelling quickly) resulting in duplication of an object.
	\item The server an object leaves may be unaware of when it should stop simulating the object (resulting in duplicated objects \cite{High-performanceServerSystems}).
	\item An object may be colliding with another object across the boundary and the collision would be missed if both objects are hosted on different servers (objects may pass through each other).
\end{itemize}

@@

There would be a trade-off with this approach, either:
\begin{enumerate}
	\item Frequent messages are sent with small corrections for objects, appearing as jitter. This would require lots of computational and network overhead. 
	\item Infrequent messages with large corrections for objects, appearing as objects moving large distances instantaneously but would require less computational and network overhead.
\end{enumerate}

@@

This approach would be affected by network delays. For example, using the stacked cubes example. Say the two cubes are free-falling and the bottom cube is being simulated on a different server, the position of the bottom cube would only be updated whenever a message was received from the remote server. If there was a long delay between messages from the remote server, for example, from a spike in latency or a sudden large loss in packets, the cube would not be updated and appear to have stopped falling. As the bottom cube has stopped on the receiving server, the top cube remains stationary, as it is in contact with the bottom cube. On the remote server, if messages are being received, the cube would be stationary and appear to be floating, as from its server's perspective, it is in contact with the bottom cube. As a result, both servers have incorrect states compared to a centralised simulation.

\subsection{All object interactions on the same server}
Simulate all objects that are interacting with each other on the same server. Auras could be used to determine which objects are potentially interacting. Auras of objects would need to be exchanged. Solution would need to manage the number of auras exchanged to reduce this overhead.

@@

In AP, an object maintains an aura, an area of interest around the object, that indicates its possible future location and, therefore, aids in identifying near future interactions. In other words, auras act as a request for objects to be migrated. Any object colliding with the aura is migrated to the server the aura came from. The presence of an aura allows: 
\begin{itemize}
	\item The prediction of future hosting requirements allowing time to transfer objects.
	\item A narrowing of interest in only considering a subset of objects promoting a scalable solution.
	\item An unhindered physics simulation for existing objects.
	\item Limiting communications requirements between servers based on focusing message passing overhead on interacting objects (promoting scalability).
\end{itemize}
In essence, AP ensures that any two objects that may be interacting are always being simulated on the same server. For example, objects C and D illustrated in Fig. \ref{AuraProj}. Object C projects its aura into Server 1, Object D collides with the aura and is migrated to Server 0, allowing continued interaction to occur.

@@

\section{Aura Calculation}
Calculating an aura size is key to creating a scalable solution balanced against consistency:
\begin{enumerate}
	\item Auras that are too small will result in a faster simulation but with more missed interactions.
	\item Auras that are too large will include more calculations that are simply not required, increasing server load and decreasing scalability.
\end{enumerate}
Therefore, how auras are calculated will be carefully described and justified.

@@

The solution to this problem is to dynamically move the server-region boundaries to prevent clusters of objects having the majority of the objects outside of their host server region or the server-region boundaries could be moved in such a way as to avoid intersecting large clusters of objects entirely. However, this solution is not yet implemented in AP.

@@

\section{Problem Definition Summary}
In this chapter, the following points have been addressed:
\begin{itemize}
	\item The requirements of a solution to scalable real-time physics.
	\item The challenges of distributing a real-time physics engine and potential solutions.
	\item A description and justification of AP, including the aura calculation.
	\item The implemented solutions to the problems of: 3-object thrashing; islands; and the corner case.
	\item The proposed solutions to: sub-optimal object hosting and 2-object thrashing, which are not yet implemented.
\end{itemize}

### 2020-03-24 ChapterResults/chapterResults.tex
Updated: Total proof-read and conclusion changed to approximate solution

It will now be explained how a naive system can lead to erroneous collision behaviour and AP can lead to erroneous collision behaviour if speed, frame-time, or latency exceed the user-defined tolerances.

@@

\subsection{Collision Error Detection}
It is possible to detect late collisions between objects; Given the relative speed of the two objects and the physics time-step, it possible to calculate the maximum expected penetration distance. If a collision is detected and it is above this value, it means it is a late collision. In addition, if no collisions occur, it means there is a missed collision.

@@

In all the following experiments collisions were recorded, for each collision, the following data is contained: Relative Speed, the speed of the two colliding objects relative to each other; Penetration, the distance the two objects overlapped when the collision was detected. Note that PhysX reports penetration of collisions as negative values and can report collisions before the two objects are intersecting, which is recorded as a positive penetration. In all experiments, the penetration is negated for simplicity. The relative speed and penetration can be used to determine if the collision was a late collision. In all the following experiment scenarios only two objects are used, meaning if there was a missed collision there would be no collision output and multiple collisions from the same pair of objects can be ignored. The two objects start the scenario with a specified velocity (drag and gravity are disabled), the direction being directly towards the other object and the two objects were given the same speed and a random variation of between $0$ and $-1m\mathord{\cdot}s^{-1}$ to prevent identical results between iterations. Spheres were used, so rotational effects, such as rotational velocity, do not need to be accounted for.

Thrashing between two objects on different servers is an unsolved problem in AP; thrashing occurrences can be detected by counting the number of migrations that occur before the first collision is detected, as only two objects are involved in each experiment, the number of migrations before a collision should only be one, otherwise thrashing has occurred. For the purposes of these experiments, collisions involving thrashing will be excluded, regardless of the correctness of the collision. Note, for the remainder of this chapter, thrashing refers to only thrashing between two objects and not the solved 3-object-thrashing problem.

@@

In order t rigorously test the hypothesis "Can real-time physics simulations remain correct when scaled?", a `worst-case' scenario was used for the experiments, in which two objects are moving directly towards each other and collide while one of the objects is intersecting the boundary between servers. This experiment setup is illustrated in Fig. \ref{fig_CorrectnessExperiment}. In this experiment two servers were used with one sphere being created on each server. The two spheres were given starting positions so the two would collide at a point when one of the spheres is intersecting the server region boundary, thus creating the most likely situation for spheres collide with each other in the first time-step after being migrated; this is necessary as AP can only lead to late or missed collisions when at least one sphere involved in the collision has just migrated (received since the last physics time-step). This is because non-migrating spheres behave the same as spheres in a centralised solution, therefore only collisions that involve a sphere that has just migrated are considered in the experiment results.

@@

\subsubsection{Evaluation - Frame-Time}
As frame-time increases, the proportion of late collisions to correct collisions increases, as shown in Figs. \ref{fig_RatioVsLatencyLow} and \ref{fig_RatioVsLatencyMid}, however in \ref{fig_RatioVsLatencyHigh} the sample size is small above the tolerance line due to the high occurrences of thrashing and so the relationship in this experiment can not be deduced. For the $8ms$ and $15ms$ experiments the late collisions do not exceed 10\% and is assumed to be accounted for by false-positives. In the $8ms$ frame-time tolerance experiment, late collisions are not expected to occur until $16ms$ frame-time, based on the aura size. Late collisions do not exceed $20\%$ and is assumed to be accounted for by false-positives. In the $15ms$ frame-time tolerance experiment, late collisions exceed $20\%$ once $16ms$ latency is exceeded as predicted by the aura size and increase to over $75\%$ at $30ms$ latency.
In the final frame-time tolerance experiment using $32ms$ frame-time tolerance, late collisions only occur for frame-times of $64ms$. The aura calculation predicts that late collisions should begin to occur as soon as $32ms$ frame-time is exceeded, however, due to the low sample size, as a result of the high occurrences of thrashing, the plot does not accurately represent any relationship between frame-time and proportion of late collisions.

For the first two frame-time experiments ($8.33ms$ and $15ms$ tolerance values), as frame-time increases, the mean penetration time increases and standard deviation also increases, shown in Figs. \ref{fig_CollisionsPenVsFrameTimeLow} and \ref{fig_CollisionsPenVsFrameTimeMid}. Standard deviation increases as the delay between frame-time update and physics update can always vary between 0 and the frame-time, meaning even in high frame-time situations, the delay between frame-time update and physics update can be 0. Exceeding the frame-time tolerance very quickly leads to high max error, this is due to the delay in messages being exchanged being created by the long frame-times on both servers (this is reflected in the aura calculation by the frame-time tolerance being multiplied by 2). However, as the target frame-time increases over the tolerance, variation in error increases and maximum errors become less likely. For the final frame-time experiment, ($33.33ms$ tolerance value), thrashing occurs the majority of the time both above (as demonstrated in \ref{tab_BelowTolerance} and \ref{tab_AboveTolerance}) and below the tolerance, due to this the results shown in \ref{fig_CollisionsPenVsFrameTimeHigh} have a very small sample size and thus will not accurately represent the relationship between frame-time and penetration.

\subsubsection{Varying Factors - Summary}
The varying factors experiments demonstrate that the aura calculation is conclusively correct for speed and frame-time aura tolerance factors and at minimum approximate for latency. For each varied factor, for all three tolerance values chosen, below the tolerance value, the mean + two standard deviations lies under the maximum penetration time expected on a centralised simulation, with only a small number of exceptions discussed above. These experiments also demonstrate the relationship between increasing the tolerance values and amount of thrashing expected to occur. There are no missed collisions, with the exception of the speed experiments, which have significantly fewer than $1\%$ misses below the tolerance and $1.12\%$ misses above the tolerance. Once factors increase above their respective tolerance values, late collisions begin to occur and the more the factors are exceeded, the number and magnitude of errors increase. This demonstrates the correctness of the aura calculation in AP.
%Therefore, we have shown that if the aura calculation factors remain within their respective tolerances, erroneous collisions will be small in magnitude and rare in number. %TODO: Summary Table: Row for each factor: Number of collisions below tolerance, Percentage of errors below tolerance, Mean error below tolerance

>>>>

A recap of how a naive system can lead to erroneous collision behaviour and how AP can lead to erroneous collision behaviour if speed, frame-time, or latency exceed the user-defined tolerances will now be given.

@@

\subsection{Collision Error Detection}
It is possible to detect late collisions between objects; Given the relative speed of the two objects and the physics time-step, it is possible to calculate the maximum expected penetration distance. If a collision is detected and it is above this value, it means it is a late collision. In addition, if no collisions occur, it means there is a missed collision.

@@

In all the following experiments collisions were recorded, for each collision, the following data is contained: Relative Speed, the speed of the two colliding objects relative to each other; Penetration, the distance the two objects overlapped when the collision was detected. PhysX reports penetration of collisions as negative values and can report collisions before the two objects are intersecting, which is recorded as a positive penetration. In all experiments, the penetration is negated for simplicity. The relative speed and penetration can be used to determine if the collision was a late collision. In all the following experimental scenarios only two objects are used, meaning if there was a missed collision there would be no collision output and multiple collisions from the same pair of objects can be ignored. The two objects start the scenario with a specified velocity (drag and gravity are disabled), the direction being directly towards the other object and the two objects were given the same speed and a random variation of between $0$ and $-1m\mathord{\cdot}s^{-1}$ to prevent identical results between iterations. Spheres were used, so rotational effects, such as rotational velocity, do not need to be accounted for.

Thrashing between two objects on different servers is an unsolved problem in AP; thrashing occurrences can be detected by counting the number of migrations that occur before the first collision is detected, as only two objects are involved in each experiment, the number of migrations before a collision should only be one, otherwise thrashing has occurred. For the purposes of these experiments, collisions involving thrashing will be excluded, regardless of the correctness of the collision. For the remainder of this chapter, thrashing refers to only thrashing between two objects and not the solved 3-object-thrashing problem.

@@

In order to rigorously test the hypothesis "Can real-time physics simulations remain correct when scaled?", a `worst-case' scenario was used for the experiments, in which two objects are moving directly towards each other and collide while one of the objects is intersecting the boundary between servers. This experiment setup is illustrated in Fig. \ref{fig_CorrectnessExperiment}. In this experiment two servers were used with one sphere being created on each server. The two spheres were given starting positions so the two would collide at a point when one of the spheres is intersecting the server region boundary, thus creating the most likely situation for spheres collide with each other in the first time-step after being migrated; this is necessary as AP can only lead to late or missed collisions when at least one sphere involved in the collision has just migrated (received since the last physics time-step). This is because non-migrating spheres behave the same as spheres in a centralised solution, therefore only collisions that involve a sphere that has just migrated are considered in the experiment results.

@@

\subsubsection{Evaluation - Frame-Time}
As frame-time increases, the proportion of late collisions to correct collisions increases, as shown in Figs. \ref{fig_RatioVsLatencyLow} and \ref{fig_RatioVsLatencyMid}, however in \ref{fig_RatioVsLatencyHigh} the sample size is small above the tolerance line due to the high occurrences of thrashing and so the relationship in this experiment can not be deduced. For the $8ms$ and $15ms$ experiments the late collisions do not exceed 10\% and are assumed to be accounted for by false-positives. In the $8ms$ frame-time tolerance experiment, late collisions are not expected to occur until $16ms$ frame-time, based on the aura size. Late collisions do not exceed $20\%$ and is assumed to be accounted for by false-positives. In the $15ms$ frame-time tolerance experiment, late collisions exceed $20\%$ once $16ms$ latency is exceeded as predicted by the aura size and increase to over $75\%$ at $30ms$ latency.
In the final frame-time tolerance experiment using $32ms$ frame-time tolerance, late collisions only occur for frame-times of $64ms$. The aura calculation predicts that late collisions should begin to occur as soon as $32ms$ frame-time is exceeded, however, due to the low sample size, as a result of the high occurrences of thrashing, the plot does not accurately represent any relationship between frame-time and proportion of late collisions.

For the first two frame-time experiments ($8.33ms$ and $15ms$ tolerance values), as frame-time increases, the mean penetration time increases and standard deviation also increases, shown in Figs. \ref{fig_CollisionsPenVsFrameTimeLow} and \ref{fig_CollisionsPenVsFrameTimeMid}. Standard deviation increases as the delay between frame-time update and physics update can always vary between 0 and the frame-time, meaning even in high frame-time situations, the delay between frame-time update and physics update can be 0. Exceeding the frame-time tolerance very quickly leads to high max error. This is due to the delay in messages being exchanged being created by the long frame-times on both servers (this is reflected in the aura calculation by the frame-time tolerance being multiplied by 2). However, as the target frame-time increases over the tolerance, variation in error increases and maximum errors become less likely. For the final frame-time experiment, ($33.33ms$ tolerance value), thrashing occurs the majority of the time (as demonstrated in \ref{tab_BelowTolerance} and \ref{tab_AboveTolerance}), both above and below the tolerance, due to this the results shown in \ref{fig_CollisionsPenVsFrameTimeHigh} have a very small sample size and thus will not accurately represent the relationship between frame-time and penetration.

\subsubsection{Varying Factors - Summary}
The varying factors experiments demonstrate that the aura calculation is conclusively correct for speed and frame-time aura tolerance factors and at minimum approximate for latency. For each varied factor, for all three tolerance values chosen, below the tolerance value, the mean + two standard deviations lies under the maximum penetration time expected on a centralised simulation, with only a small number of exceptions discussed above. These experiments also demonstrate the relationship between increasing the tolerance values and amount of thrashing expected to occur. There are no missed collisions, with the exception of the speed experiments and these have significantly fewer than $1\%$ misses below the tolerance and $1.12\%$ misses above the tolerance. Once factors increase above their respective tolerance values, late collisions begin to occur and the more the factors are exceeded, the number and magnitude of errors increase. This demonstrates the correctness of the aura calculation in AP.
%Therefore, we have shown that if the aura calculation factors remain within their respective tolerances, erroneous collisions will be small in magnitude and rare in number. %TODO: Summary Table: Row for each factor: Number of collisions below tolerance, Percentage of errors below tolerance, Mean error below tolerance
